---
title: "Geodatabase QA Output Summary Script"
author: "Andrew Brown (andrew.g.brown@ca.usda.gov)"
version: "0.1"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
### Purpose:
This report summarizes contents of QA shapefiles from SSURGO QA Toolbox. Tabular summaries will work with any `QA_*.shp` files containing a field matching a user-specified MUSUM column name (default: "MUSYM"). The table shows the summarized contents of the a status field (default name: "Status"; default levels: blank, "OK", and "Error"). Output is written back to the QA file path as CSV named `~status_summary_*.csv`.

__Columns of tabular output are:__

*  MUSYM
*  number features per MUSYM
*  number features as Error
*  number of features Unchecked
          
____

```{r, echo=F, message=F, warning=F, results='asis'}
#path to QA Toolbox shapefile output
#qafilepath <- "S:/NRCS/Archive_Andrew_Brown/CA630/QA/CommonSoilLines/20180110/"
qafilepath <- "L:/NRCS/MLRAShared/CA630/FG_CA630_OFFICIAL.gdb"
musym.column.name <- "MUSYM" #column name containing MUSYM (the variable you are stratifying on)
polyid.column.name <- "POLYID"
areasym.column.name <- "AREASYMBOL" #column name containing MUSYM (the variable you are stratifying on)
status.column.name <- "Status" #column name containing status information you would like to summarize by MUSYM
ok.pattern <- "OK" #pattern to match for records with "fixed" status
error.pattern <- "Error" #pattern to match records that have been reviewed and determined to be an error needing further attention
```

```{r, echo=F, message=F, warning=F, results='asis'}
library(rgdal)
library(knitr)
gdb.layers <- ogrListLayers(dsn=qafilepath)

keep.qa <- grepl(gdb.layers,pattern="^QA_")
gdb.layers.qa <-  gdb.layers[keep.qa]

keep._a <-  grepl(gdb.layers, pattern="_a")
spatialsrc <- gdb.layers[keep._a]

filedataframes <- list()
for(q in gdb.layers.qa) {
  filedataframes[[q]] <- try(readOGR(dsn=qafilepath, layer=q))
}

sp_a <- readOGR(dsn=qafilepath, layer=spatialsrc)
sp_a@data$POLYID_i <- sapply(slot(sp_a, "polygons"), function(x) slot(x, "ID")) 

#Summarizes QA output by mapunit symbol
for(f in 1:length(filedataframes)) {
  fil <- filedataframes[[f]]
  musymcol <- which(grepl(names(fil), pattern=musym.column.name, ignore.case = T))
  if(length(musymcol) <= 0) { #ONLY CREATES OUTPUT IF THERE IS A MUSYM FIELD TO STRATIFY ON. so, if missing, make one from joining on polyid to _a
    polyidcol <- which(grepl(names(fil), pattern=polyid.column.name, ignore.case = T))
    if(length(polyidcol)) { 
      fil@data[[musym.column.name]] <- merge(fil@data, sp_a@data, by.x="POLYID", by.y="POLYID_i", all.x=T)$MUSYM
      musymcol <- which(grepl(names(fil), pattern=musym.column.name, ignore.case = T))
      #writeOGR(fil, dsn = qafilepath, layer=gdb.layers.qa[[f]], driver="OpenFileGDB")
    } else {
      print(paste0("Could not locate MUSYM or POLYID field (to infer MUSYM) for layer: ", gdb.layers.qa[[f]]))
      next
    }
  }
  cat(paste0("\n#### ", gdb.layers.qa[[f]]),"\n")
  if(length(musymcol) > 1) #for case e.g. common lines where there is a L_MUSYM & R_MUSYM, just take the first column matching pattern
    musymcol <- musymcol[1]
  df <- aggregate(fil@data[,musymcol],by=list(fil@data[,musymcol]),FUN=length)
  df <- cbind(df,aggregate(fil@data[status.column.name],by=list(fil@data[,musymcol]),FUN=function(x) return(sum(as.numeric(grepl(x, pattern=ok.pattern)))))[,2])
  df <- cbind(df,aggregate(fil@data[status.column.name],by=list(fil@data[,musymcol]),FUN=function(x) return(sum(as.numeric(grepl(x, pattern=error.pattern)))))[,2])
  names(df) <- c("MUSYM", "Features", ok.pattern, error.pattern)
  df <- transform(df, Unchecked = Features - OK - Error)
  df$Unchecked <- df$Features - df[ ,ok.pattern] - df[ ,error.pattern]
  print(kable(df))
  write.csv(df,file=paste0(qafilepath,"\\~status_summary_",strsplit(gdb.layers.qa[[f]],"\\.")[[1]][1],".csv"))
}
```
