---
title: "Geodatabase QA Output Summary Script"
author: "Andrew Brown (andrew.g.brown@ca.usda.gov)"
version: "0.1"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
### Purpose:
This report summarizes contents of QA shapefiles from SSURGO QA Toolbox. Tabular summaries will work with any `QA_*.shp` files containing a field matching a user-specified MUSUM column name (default: "MUSYM"). The table shows the summarized contents of the a status field (default name: "Status"; default levels: blank, "OK", and "Error"). Output is written back to the QA file path as CSV named `~status_summary_*.csv`.

__Columns of tabular output are:__

*  MUSYM
*  number features per MUSYM
*  number features as OK
*  number features as Error
*  number of features Unchecked
          
____

```{r, echo=F, message=F, warning=F, results='asis'}
#path to QA Toolbox shapefile output
qafilepath <- "S:\\NRCS\\Archive_Andrew_Brown\\CA630\\QA\\CommonSoilLines\\20171120"
musym.column.name <- "MUSYM" #column name containing MUSYM (the variable you are stratifying on)
status.column.name <- "Status" #column name containing status information you would like to summarize by MUSYM
ok.pattern <- "OK" #pattern to match for records with "fixed" status
error.pattern <- "Error" #pattern to match records that have been reviewed and determined to be an error needing further attention
```

```{r, echo=F, message=F, warning=F, results='asis'}
library(rgdal)
library(knitr)
qafilepath <- chartr('\\','/',qafilepath) 
qafilesnopath <- list.files(qafilepath)
keep <- grepl(qafilesnopath,pattern="QA_.*shp$")
qafilesnopath <- qafilesnopath[keep]
qafiles <- paste0(qafilepath,"/",qafilesnopath)
filedataframes <- lapply(qafiles,FUN=function(x) return(readOGR(dsn=x, verbose=F)@data))

#Summarizes QA output by mapunit symbol
for(f in 1:length(filedataframes)) {
  fil <- filedataframes[[f]]
  musymcol <- which(grepl(names(fil), pattern=musym.column.name, ignore.case = T))
  if(length(musymcol) > 0) { #ONLY CREATES OUTPUT IF THERE IS A MUSYM FIELD TO STRATIFY ON!
    cat(paste0("\n#### ", qafiles[[f]]),"\n")
    musymcol <- musymcol[1]
    df <- aggregate(fil[,musymcol],by=list(fil[,musymcol]),FUN=length)
    df <- cbind(df,aggregate(fil[status.column.name],by=list(fil[,musymcol]),FUN=function(x) return(sum(as.numeric(grepl(x, pattern=ok.pattern)))))[,2])
    df <- cbind(df,aggregate(fil[status.column.name],by=list(fil[,musymcol]),FUN=function(x) return(sum(as.numeric(grepl(x, pattern=error.pattern)))))[,2])
    names(df) <- c("MUSYM", "Features", ok.pattern, error.pattern)
    df <- transform(df, Unchecked = Features - OK - Error)
    df$Unchecked <- df$Features - df[ ,ok.pattern] - df[ ,error.pattern]
    print(kable(df))
    write.csv(df,file=paste0(qafilepath,"\\~status_summary_",strsplit(qafilesnopath[[f]],"\\.")[[1]][1],".csv"))
  } else {
    print(paste0("Unable to locate MUSUM field (name: ", musum.column.name))
  }
}
```
