---
title: "RV Pedon Overlay"
version:  "0.1b"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
#### Instructions:
1. Load all possible component pedons into NASIS selected set
2. Load DMUs of interest into NASIS selected set
3. Ensure polygon data source name and layer are correct and update config as needed.

#### Output:
Table containing DMU MUSYM, component name, RV component pedon ID, RV component pedon MUSYM. Last two columns check DMU v.s. Pedon match and Borrowed Pedon Note, respectively.

```{r config}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE)

poly.dsn = "L:/NRCS/MLRAShared/CA630/FG_CA630_OFFICIAL.gdb"
poly.layer = "ca630_a"
dmudescprefix = "CA630" #usually survey area, the portion of dmudesc that is NOT the musym from the spatial
rv_only = TRUE #filter component pedons for rvindicator = 1; note: script does not support rv_only=FALSE
```

```{r setup, include=FALSE}
library(rgdal)
library(plyr)
library(soilDB)
library(knitr)

f <- fetchNASIS() #load point data (need to have all potential copedons in selected set)
mu <- readOGR(dsn = poly.dsn, layer = poly.layer, stringsAsFactors=FALSE) # load spatial

get_NASIS_componentpedons <- function(rv_only=FALSE) {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT copedoniid, coiidref, peiidref, rvindicator, recwlupdated, recuseriidref FROM copedon;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  if(rv_only)
    d<-d[which(d$rvindicator == 1),]
  return(d)
}

get_NASIS_componenttext <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT cotextiid, coiidref, recdate, recauthor, comptextkind, textcat, textsubcat, textentry, recwlupdated, recuseriidref FROM cotext;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

good.idx <- which(!is.na(f$x_std))
f.bad <- f[-good.idx, ]
f <- f[good.idx, ]
coordinates(f) <- ~ x_std + y_std
proj4string(f) <- '+proj=longlat +datum=WGS84'
f.sub <- as(f,'SpatialPointsDataFrame')
f.sub <- spTransform(f.sub, CRS(proj4string(mu)))
f$pMUSYM <- over(f.sub, mu)$MUSYM

sitexcomponent <- merge(site(f), get_NASIS_componentpedons(rv_only=T), by.x='peiid', by.y='peiidref', all=T) #join copedon to pedon table, keep only records that are copedons  

componentz <- fetchNASIS_components()
componentz <- merge(site(componentz), sitexcomponent, by.x = 'coiid', by.y='coiidref')
cotext <- get_NASIS_componenttext()
cotext.borrowed <- cotext[which(cotext$textcat == "rep pedon" & cotext$textsubcat == "borrowed"),]$coiid #list of coiids with rep pedon borrowed note
```


```{r}
componentz$dmuid <- sub(componentz$dmudesc, pattern = dmudescprefix, replacement = "")
unitmatch <- componentz$pMUSYM==componentz$dmuid
componentz$has_note = F
componentz[!unitmatch, ]$has_note <- (componentz[!unitmatch, ]$coiid %in% cotext.borrowed)
df <- data.frame(componentz$dmuid, componentz$compname, componentz$pedon_id, componentz$pMUSYM, unitmatch, (!unitmatch & !componentz$has_note))
df <- df[order(componentz$dmuid, componentz$compname),]
names(df) <- c("DMU","Component Name", "PedonID", "PedonMUSYM", "Match", "Needs borrowed note?")
kable(df, row.names = F)
```