---
title: "Pedon Mapunit and Area Overlap Report"
author: "Andrew Brown"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r,echo=F}
library(rgdal)
library(soilDB)
library(pacman)

mu <- readOGR(dsn = 'L:/NRCS/MLRAShared/CA630/FG_CA630_OFFICIAL.gdb', layer = 'ca630_a', stringsAsFactors=FALSE)
surveybdy <- readOGR(dsn = 'L:/NRCS/MLRAShared/CA630/FG_CA630_OFFICIAL.gdb', layer = 'ca630_b', stringsAsFactors=FALSE)
mlrabdy <- readOGR(dsn = 'F:/Geodata/mlra', layer = 'mlra_z10_a_mo2', stringsAsFactors=F)

surveybdy$areasymbol <- "CA630"

x <- fetchNASIS()
x$taxonname = toupper(x$taxonname)

good.idx <- which(!is.na(x$x_std))
x.bad <- x[-good.idx, ]
x <- x[good.idx, ]
coordinates(x) <- ~ x_std + y_std
proj4string(x) <- '+proj=longlat +datum=NAD83'
x.sub <- as(x, 'SpatialPointsDataFrame')
x.sub <- spTransform(x.sub, CRS(proj4string(mu)))
site(x)$musym <- over(x.sub, mu)$MUSYM 
x.sub <- spTransform(x.sub, CRS(proj4string(surveybdy)))
site(x)$areasym[!is.na(over(x.sub, surveybdy)[,1])] <- surveybdy$areasymbol
x.sub <- spTransform(x.sub, CRS(proj4string(mlrabdy)))
site(x)$mlra <-over(x.sub, mlrabdy)$MLRA2002
```

```{r, echo=F}
get_site_area_info_from_NASIS_db <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT siteiid as siteiid, usiteid, stateareaiidref, stateareatypeiidref, countyareaiidref, countyareatypeiidref, mlraareaiidref, mlraareatypeiidref FROM site;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

get_site_area_overlap_from_NASIS_db <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT siteiidref as sareaoviid, siteiidref, areaiidref, recwlupdated, recuseriidref FROM siteaoverlap;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

get_site_mu_overlap_from_NASIS_db <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT smuoviid as smuoviid, siteiidref, lmapunitiidref, recwlupdated, recuseriidref FROM sitemuoverlap;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

get_legend_mu_from_NASIS_db <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT lmapunitiid as liidref, lmapunitiid, musym, muiidref, mustatus, muacres, farmlndcl, farmlndclsec, recwlupdated, recuseriidref FROM lmapunit;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

get_area_NASIS_db <- function() {
  if(!requireNamespace('RODBC'))
    stop('please install the `RODBC` package', call.=FALSE)
  q <- "SELECT areatypeiidref as areatypeiidref, areaiid, areasymbol, areaname, areaacres, obterm FROM area;"
  channel <- RODBC::odbcDriverConnect(connection="DSN=nasis_local;UID=NasisSqlRO;PWD=nasisRe@d0n1y")
  d <- RODBC::sqlQuery(channel, q, stringsAsFactors=FALSE)
  RODBC::odbcClose(channel)
  return(d)
}

site_info <- get_site_area_info_from_NASIS_db()
site_area_overlap <- get_site_area_overlap_from_NASIS_db()
site_mu_overlap <- get_site_mu_overlap_from_NASIS_db()
legend_mu <- get_legend_mu_from_NASIS_db()
area_info <- get_area_NASIS_db()

area_sym_list = area_name_list = area_acres_list = list()
area_sym_list[area_info$areaiid] = area_info$areasymbol
area_name_list[area_info$areaiid] = area_info$areaname
area_acres_list[area_info$areaiid] = area_info$areaacres

site <- merge(site_info,site_area_overlap,by.x="siteiid",by.y="siteiidref")

site$stateareasym <- area_sym_list[site$stateareaiidref]
site$stateareaname <- area_name_list[site$stateareaiidref]
site$stateareaacres <- area_acres_list[site$stateareaiidref]

site$countyareasym <- area_sym_list[site$countyareaiidref]
site$countyareaname <- area_name_list[site$countyareaiidref]
site$countyareaacres <- area_acres_list[site$countyareaiidref]

site$mlraareasym <- area_sym_list[site$mlraareaiidref]
site$mlraareaname <- area_name_list[site$mlraareaiidref]

site$areaoverlapsym <- area_sym_list[site$areaiidref]
site$areaoverlapname <- area_name_list[site$areaiidref]
site_area = site #site area contains the data (usually multople areas per site) from AREA overlap table

df_site_area <- aggregate(cbind(site_area$usiteid, site_area$areaoverlapsym),by = list(site_area$usiteid), FUN=paste)[,c(1,3)]
names(df_site_area) <- c("USERSITEID", "AREAOVERLAP")

site_mu = merge(site_info[,1:2],site_mu_overlap,by.x="siteiid",by.y="siteiidref")
site_mu = merge(site_mu, legend_mu, by.x="lmapunitiidref", by.y="lmapunitiid")

# S <- site[site$usiteid == "06CA630JCR002",]
# smu <- site_mu[site_mu$usiteid == "06CA630JCR002",] ### sanity check!
```

```{r}
#lets consolidate...
gis_names = c("peiid", "pedon_id", "siteiid", "site_id", "obs_date", "utmzone", "utmeasting" , "utmnorthing" ,"x","y","datum","musym","areasym","mlra")
nasis_areao_names = c("siteiid","usiteid","stateareasym","countyareasym","mlraareasym","areaoverlapsym")
nasis_muo_names = c("siteiid","usiteid","musym")

gis <- site(x)[,gis_names]
areao <- site_area[,nasis_areao_names]
areao_flat <- df_site_area
muo <- site_mu[,nasis_muo_names]
nrow(gis) #number of records with valid spatial data WITH a pedon record... can we just pull site data?
nrow(areao_flat) #number of records with one or more areaoverlap records (flattened to one record using agg->paste)
nrow(muo) #number of pedons in nasis with mapunit overlap set

#most serious order of business... which pedons have been assigned a MU in muoverlap... but are no longer in that mapunit!
gis$musym_match <- NA
compare.idx <- which(gis$site_id %in% muo$usiteid)
for(c in compare.idx) {
  rec = muo[which(gis[c,]$site_id == muo$usiteid),]
  if(is.na(gis[c,]$musym)) {
    gis[c,]$musym_match = -1
  } else if(any(gis[c,]$musym != rec$musym)) {
    gis[c,]$musym_match = 0
  } else {
    gis[c,]$musym_match = 1
  }
}

#list of pedons with incorrect muoverlap
gis_inc <- gis[which(gis$musym_match != 1),]
write.csv(file="sites_incorrect_muoverlap.csv",x=gis_inc)

#list of sites with NO muoverlap
gis_nomuo <-  gis[-compare.idx,]
write.csv(file="sites_no_muoverlap.csv",x=gis_nomuo)

#this spits out the number of muoverlap records that are -1:nogiscomparison;0:incorrect;1:correct
aggregate(gis$musym_match,by=list(gis$musym_match),FUN=length)

#this spits out any user site ids with multiple MUs assigned in MUoverlap 
muo[which(aggregate(muo$musym, by=list(muo$siteiid), FUN=length)[,2] > 1),]$usiteid

#this spits out potential copied site records (i.e. same user site id)
muo[which(aggregate(muo$musym, by=list(muo$usiteid), FUN=length)[,2] > 1),]$usiteid
