---
title: "Lab Pedon Check"
author: "Andrew Brown"
output: html_document
---

```{r setup, include=FALSE}
library(soilDB)
library(knitr)

knitr::opts_chunk$set(echo = FALSE)

#lab_pedons <- read.csv("S:/NRCS/Archive_Andrew_Brown/CA630/QC/LabPedons/ca630_lab_pedons.csv", stringsAsFactors = FALSE)
lab_pedons <- read.csv("S:/NRCS/Archive_Andrew_Brown/CA630/QC/LabPedons/ca630_lab_pedons_fixed.csv", stringsAsFactors = FALSE)

nasis_pedons <- fetchNASIS()
```
### Pre-cleaning

A few pedons were causing problems. 

#### Two SoilVeg pedons had incorrect overlap (now fixed & can be queried  PEDON/SITE by SSA ID, SSA='CA630')
 - 67-CA-05-042x - previously assigned CA628
 - 75-CA-55-024x - previously assigned CA731

#### One had incorrect coordinates; it is not actually in CA630
 - S1958CA039001 - Site description says MLRA 17 granitic alluvium, Madera area. NASIS point is near Tuttletown east of New Melones off HWY 49. Likely that Lat/Lng coordinates are the error, as they don't match the Township and Range description.

#### Two pedon records with no horizon data, but are located in CA630
 - S04-hwy49-719
 - S04-hwy49-720


## List (sorted by USERPEDONID) of taxon names - non-matching to be updated on official list
```{r}
# List of lab pedons from official list lacking a matching userpedonid in NASIS (these are likely overlap or other types of errors, note fetchNASIS pedon QC)
#missing.in.nasis <- data.frame(userpedonid=lab_pedons[!(lab_pedons$userpedonid %in% nasis_pedons$pedon_id),]$userpedonid)
# kable(missing.in.nasis)
# 
lab_nasis_pedons <- nasis_pedons[nasis_pedons$pedon_id %in% lab_pedons$userpedonid,]
lab_nasis_pedons <- site(lab_nasis_pedons[order(lab_nasis_pedons$pedon_id),])
lab_pedons <- lab_pedons[order(lab_pedons$userpedonid),]
lab.pedon.compare <- merge(lab_pedons, lab_nasis_pedons, by.x="userpedonid", by.y="pedon_id")
test.frame <- data.frame(pedonid=lab.pedon.compare$userpedonid, SOURCE_NAME=lab.pedon.compare$series, NASIS_NAME=lab.pedon.compare$taxonname)
test.frame$match <- (as.character(lab.pedon.compare$series) == as.character(lab.pedon.compare$taxonname))
kable(test.frame)
```

## List of problem pedons (should be removed from list altogether)
```{r}
kable(data.frame(userpedonid=lab_pedons[!(lab_pedons$userpedonid %in% lab.pedon.compare$userpedonid),]$userpedonid))
```


