---
title: "Carbonates in Dark Surface Epipedons"
author: "Andrew G. Brown"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", echo = FALSE, message = FALSE)
Sys.setenv(TAR_MAKE_REPORTER = "silent")
```

```{r}
library(aqp)
library(targets)

tar_unscript()

hzkablefun <-
  function(spc) {
    knitr::kable(horizons(spc)[, c(
      "labsampnum",
      "phiid",
      "peiid",
      "hzname",
      "hzdept",
      "hzdepb",
      'caco3_lt_2_mm',
      'd_value','m_value','m_chroma',
      'carbon_combined'
    )])
  }
```

```{targets mollic_umbric_lab_morph}
tar_option_set(packages = c("magrittr","aqp"))
list(
  tar_target(name = mollic_umbric_lab_morph,
             read.csv(file = "mollic_umbric_lab_morph.csv", stringsAsFactors = FALSE)
  ), 
  tar_target(name = mollic_umbric_lab_morph_complete,{
             ldx <- complete.cases(mollic_umbric_lab_morph[,c('hzdept','caco3_lt_2_mm','isDark','carbon_combined')])
             mollic_umbric_lab_morph_complete <- mollic_umbric_lab_morph[ldx, ]
             depths(mollic_umbric_lab_morph_complete) <- peiid ~ hzdept + hzdepb
             mollic_umbric_lab_morph_complete$isDarkMoist <- mollic_umbric_lab_morph_complete %>% 
               hasDarkColors(d_value = NA)
             mollic_umbric_lab_morph_complete
    }),
  tar_target(name = q1_data, mollic_umbric_lab_morph_complete %>% subsetHz(carbon_combined >= 0.6 &
                                                                          caco3_lt_2_mm >= 15 & 
                                                                           !isDark | 
                                                                            (is.na(isDark) & !isDarkMoist))),
  tar_target(name = q2_data, mollic_umbric_lab_morph_complete %>% subsetHz(carbon_combined >= 0.6 &
                                                                          caco3_lt_2_mm >= 40 & 
                                                                           !isDark | 
                                                                            (is.na(isDark) & !isDarkMoist))),
  tar_target(name = q1_data_18, q1_data %>% subsetHz(hzdept <= 18)),
  tar_target(name = q2_data_18, q2_data %>% subsetHz(hzdept <= 18)),
  tar_target(name = q1_data_25, q1_data %>% subsetHz(hzdept <= 25)),
  tar_target(name = q2_data_25, q2_data %>% subsetHz(hzdept <= 25))
)
```

```{r}
tar_make()
```

```{r}
tar_load(mollic_umbric_lab_morph_complete)

n_total <- nrow(horizons(mollic_umbric_lab_morph_complete))
```

Total number of KSSL mollic epipedon subhorizon records: **`r n_total`**

 -  _ASSOCIATED WITH NASIS MORPHOLOGIC RECORDS with "complete" data_ 
 
   - horizon depths
   - measured carbonate in <2mm fraction
   - organic carbon >0.6%
   - dry and moist color

# Questions


### 1) Does the data show that there are mollic subhorizons that have >= 15% CaCO3 and fail to meet the <= 5 dry value, 3/3 moist value/chroma rule 3.a. in the original 12th ed. If so, how many that are in the upper 18 cm? upper 25 cm?   

```{r}
tar_load(q1_data)
tar_load(q1_data_18)
tar_load(q1_data_25)

n_q1 <- nrow(q1_data)
n_q1_lt18 <- nrow(q1_data_18)
n_q1_lt25 <- nrow(q1_data_25)
```

There are `r n_q1` mollic subhorizons (out of `r n_total`) that have >= 15% CaCO3, meet the organic carbon requirement, and do not meet "5-3-3" colors. This is `r round(n_q1 / n_total, 2) * 100`% of the data that are complete enough to be verified.

Of those `r n_q1` mollic subhorizons n=`r n_q1_lt18` are within the upper 18cm and n=`r n_q1_lt25` are within the upper 25cm.

#### `labsampnum` (<= 18cm)

`r hzkablefun(q1_data_18)`

#### `labsampnum` (<= 25cm)

`r hzkablefun(q1_data_25)`

### 2) Does the data show that there are mollic subhorizons that have >= 40% CaCO3, and fail to meet the <= 3/3 moist value/chroma rule 3.b. in the original 12th Ed. If so, how many that are in the upper 18 cm? upper 25 cm?

```{r}
tar_load(q2_data)
tar_load(q2_data_18)
tar_load(q2_data_25)

n_q2 <- nrow(q2_data)
n_q2_lt18 <- nrow(q2_data_18)
n_q2_lt25 <- nrow(q2_data_25)
```

There are `r n_q2` mollic subhorizons (out of `r n_total`) that have >= 40% CaCO3, meet the organic carbon requirement, and do not meet "5-3-3" colors. This is `r round(n_q2/n_total, 2) * 100`% of the data that are complete enough to be verified.

Of those `r n_q2` mollic subhorizons n=`r n_q2_lt18` are within the upper 18cm and n=`r n_q2_lt25` are within the upper 25cm.

#### `labsampnum` (<= 18cm)

`r hzkablefun(q2_data_18)`

#### `labsampnum` (<= 25cm)

`r hzkablefun(q2_data_25)`

# Workflow Network Diagram

```{r}
tar_visnetwork()
```

